---
title: "CF_CueRatio_lmm"
author: "Haiyang Jin (hjin317@aucklanduni.ac.nz)"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    number_sections: true
    toc: true
version: "1.0"
---


# Preparations
```{r setup and load the related libraries, include=FALSE}
## set chunk options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(warning = FALSE)
options(width = 1200)
# knitr::opts_chunk$set(fig.width=7, fig.asp =0.618) 
# knitr::opts_chunk$set(comment = "#") # (NA, to remove all hashes)

## load libraries
# libraries for cleaning data
library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
library(optimx)
library(emmeans)

folder_lmm <- "lmm_output"

```


## Load data 
```{r read the data file}

# list filenames
file1_list <- list.files(file.path("data", "1"), pattern = "*.xlsx", full.names = TRUE)
file2_list <- list.files(file.path("data", "2"), pattern = "*.xlsx", full.names = TRUE)

# load data
df_raw_E1 <- sapply(file1_list, read_excel, na = "NA", simplify = FALSE) %>% bind_rows(.id = "id") 
df_raw_E2 <- sapply(file2_list, read_excel, na = "NA", simplify = FALSE) %>% bind_rows(.id = "id")
  

# combine data from the two experiments
df_raw_E1_temp <- df_raw_E1 %>% 
  select(-c(SubTrial)) %>% 
  mutate(Probability = 0.5,
         targetDuration = 200, # fix a bug in exp
         Participant = Participant + 100)

df_raw_E2_temp <- df_raw_E2 %>% 
  select(-c(Age, Gender, Ethnicity, Block, FaceIndex)) %>% 
  mutate(Participant = Participant + 200,
         Probability = if_else(grepl("75TopCue", Experiment), 0.75, 0.25))

df_raw <- bind_rows(df_raw_E1_temp, df_raw_E2_temp)
  
df_raw

```

## Tidy data

```{r select certain rows and columns from the data, and calculate the Z value for reaction times}

df_tidy <- df_raw %>% 
  filter(!is.na(thisResponse)) %>%  # remove NA trials based on responses
  mutate(Cue = if_else(CuedHalf == "T", "top", "bottom"),
         Congruency = if_else(Congruency == "C", "congruent", "incongruent"),
         Alignment = if_else(Alignment == "A", "aligned", "misaligned"),
         SameDifferent = if_else(SameDifferent == "S", "same", "different"),
         Participant = as_factor(Participant),
         Cue = factor(Cue, levels = c("top", "bottom")),
         Congruency = factor(Congruency, levels = c("congruent", "incongruent")),
         Alignment = as_factor(Alignment),
         SameDifferent = factor(SameDifferent, levels = c("same", "different")),
         Probability = as_factor(Probability),
         Resp = if_else(thisResponse == "S", 1, 0),
         RT = round(reactionTime * 1000+200)) # plus the duration of target faces

```

+ Cue: top vs. bottom
+ Congruency: congruent vs. incongruent
+ Alignment: aligned vs. misaligned
+ SameDifferent: same vs. different


```{r}
# Trial numbers in each condition
df_tidy %>% 
  group_by(Participant) %>% 
  summarize(nTotal = n()) 

# For 3 participants in E1, one trail was removed due to no response recorded.
```


```{r}
# set successive difference coding for fixed effects
contrasts(df_tidy$Cue) <- MASS::contr.sdif(nlevels(df_tidy$Cue)) 
contrasts(df_tidy$Congruency) <- MASS::contr.sdif(nlevels(df_tidy$Congruency)) 
contrasts(df_tidy$Alignment) <- MASS::contr.sdif(nlevels(df_tidy$Alignment))
contrasts(df_tidy$SameDifferent) <- MASS::contr.sdif(nlevels(df_tidy$SameDifferent))

# set successive difference coding for random effects
df_lmm <- df_tidy %>% 
  mutate(
    Cue_C = if_else(Cue == "top", -.5, .5),
    Con_C = if_else(Congruency == "congruent", -.5, .5),
    Ali_C = if_else(Alignment == "aligned", -.5, .5),
    Sam_C = if_else(SameDifferent == "same", -.5, .5),
    
    Cue_Con = Cue_C * Con_C,
    Cue_Ali = Cue_C * Ali_C,
    Cue_Sam = Cue_C * Sam_C,
    Con_Ali = Con_C * Ali_C,
    Con_Sam = Con_C * Sam_C,
    Ali_Sam = Ali_C * Sam_C,
    
    Cue_Con_Ali = Cue_Con * Ali_C,
    Cue_Con_Sam = Cue_Con * Sam_C,
    Cue_Ali_Sam = Cue_Ali * Sam_C,
    Con_Ali_Sam = Con_Ali * Sam_C,
    
    Cue_Con_Ali_Sam = Cue_Con_Ali * Sam_C
  )

# save the data (for fitting model in cluster)
save(df_lmm, file = file.path("data", "df_lmm.RData"))

```


# Responses (d')
## Fitting the generalized mixed models
### The maximal model
```{r resp max}

# file_resp_max <- file.path(folder_lmm, "Resp_lmm_max.RData")
# 
# # fit the max model
# if (!file.exists(file_resp_max)) {
#   glmm_resp_max <- glmer(
#     Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability +
#       (Cue * Congruency * Alignment * SameDifferent | Participant), # Con_Ali_Sam
#     family = binomial(link = "probit"),
#     data = df_lmm,
#     control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
#                            optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
#   )
# 
#   save(glmm_resp_max, file = file_resp_max)
# } else {
#   load(file_resp_max)
# }
# 
# print(summary(glmm_resp_max), corr = FALSE)

```

### The zero-correlation-parameter model
```{r resp zcp}

file_resp_zcp <- file.path(folder_lmm, "Resp_lmm_zcp.RData")

# fit the zcp model
if (!file.exists(file_resp_zcp)) {
  glmm_resp_zcp <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Con_C + Ali_C + Sam_C + 
         Cue_Con + Cue_Ali + Cue_Sam + Con_Ali + Con_Sam + Ali_Sam +
         Cue_Con_Ali + Cue_Con_Sam + Cue_Ali_Sam + Con_Ali_Sam + 
         Cue_Con_Ali_Sam || Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_zcp, file = file_resp_zcp)
} else {
  load(file_resp_zcp)
}

print(summary(glmm_resp_zcp), corr = FALSE)

```

### The reduced model
```{r PCA analysis for resp zcp lmm}
summary(rePCA(glmm_resp_zcp))
```

```{r resp reduced}
file_resp_rdc <- file.path(folder_lmm, "Resp_lmm_rdc.RData")

# fit the zcp model
# three random effects were removed
if (!file.exists(file_resp_rdc)) {
  glmm_resp_rdc <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Con_C + Ali_C + Sam_C + 
         Cue_Con + Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + 
         Cue_Con_Sam + Cue_Ali_Sam + Con_Ali_Sam +# Cue_Con_Ali + 
         Cue_Con_Ali_Sam || Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_rdc, file = file_resp_rdc)
} else {
  load(file_resp_rdc)
}

print(summary(glmm_resp_rdc), corr = FALSE)
```


```{r resp continue to fit }
file_resp_rdc1 <- file.path(folder_lmm, "Resp_lmm_rdc1.RData")

# fit the etd1 model
if (!file.exists(file_resp_rdc1)) {
  ss <- getME(glmm_resp_rdc, c("theta","fixef"))
  glmm_resp_rdc1 <- update(
    glmm_resp_rdc, start=ss,
    control=glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
} else {
  load(file_resp_rdc1)
}

print(summary(glmm_resp_rdc1), corr = FALSE)
```

```{r resp2 reduced}
file_resp_rdc2 <- file.path(folder_lmm, "Resp_lmm_rdc2.RData")

# fit the rdc2 model
if (!file.exists(file_resp_rdc2)) {
  glmm_resp_rdc2 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Con_C + Ali_C + Sam_C + 
         Cue_Con + Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + 
         Cue_Con_Sam + Con_Ali_Sam +# Cue_Con_Ali + Cue_Ali_Sam + 
         Cue_Con_Ali_Sam || Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_rdc2, file = file_resp_rdc2)
} else {
  load(file_resp_rdc2)
}

print(summary(glmm_resp_rdc2), corr = FALSE)
```

### The extended model
```{r resp exteded model glmm_resp_rdc}
file_resp_etd <- file.path(folder_lmm, "Resp_lmm_etd.RData")

# fit the etd model
if (!file.exists(file_resp_etd)) {
  glmm_resp_etd <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Con_C + Ali_C + Sam_C + 
         Cue_Con + Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + 
         Cue_Con_Sam + Con_Ali_Sam +# Cue_Con_Ali + Cue_Ali_Sam + 
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd, file = glmm_resp_etd)
} else {
  load(file_resp_etd)
}

print(summary(glmm_resp_etd), corr = FALSE)
```

`Con_C` was removed from extended model.

```{r resp exteded1 model}
file_resp_etd1 <- file.path(folder_lmm, "Resp_lmm_etd1.RData")

# fit the etd1 model
if (!file.exists(file_resp_etd1)) {
  glmm_resp_etd1 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Ali_C + Sam_C + # Con_C + 
         Cue_Con + Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + 
         Cue_Con_Sam + Cue_Ali_Sam + # Cue_Con_Ali +Con_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd1, file = glmm_resp_etd1)
} else {
  load(file_resp_etd1)
}

print(summary(glmm_resp_etd1), corr = FALSE)
```

`Cue_Con` was removed from extended1 model.

```{r resp exteded2 model}
file_resp_etd2 <- file.path(folder_lmm, "Resp_lmm_etd2.RData")

# fit the etd2 model
if (!file.exists(file_resp_etd2)) {
  glmm_resp_etd2 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Ali_C + Sam_C + # Con_C + 
         Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + 
         Cue_Con_Sam + Cue_Ali_Sam + # Cue_Con_Ali +Con_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd2, file = glmm_resp_etd2)
} else {
  load(file_resp_etd2)
}

print(summary(glmm_resp_etd2), corr = FALSE)
```

`Ali_C` was removed.

```{r resp exteded3 model}
file_resp_etd3 <- file.path(folder_lmm, "Resp_lmm_etd3.RData")

# fit the etd3 model
if (!file.exists(file_resp_etd3)) {
  glmm_resp_etd3 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Sam_C + # Con_C + Ali_C + 
         Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + 
         Cue_Con_Sam + Cue_Ali_Sam + # Cue_Con_Ali +Con_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd3, file = glmm_resp_etd3)
} else {
  load(file_resp_etd3)
}

print(summary(glmm_resp_etd3), corr = FALSE)
```

`Cue_Ali_Sam` was removed.

```{r resp exteded4 model}
file_resp_etd4 <- file.path(folder_lmm, "Resp_lmm_etd4.RData")

# fit the etd4 model
if (!file.exists(file_resp_etd4)) {
  glmm_resp_etd4 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Sam_C + # Con_C + Ali_C + 
         Cue_Ali + Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + 
         Cue_Con_Sam +  # Cue_Con_Ali +Con_Ali_Sam + Cue_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd4, file = glmm_resp_etd4)
} else {
  load(file_resp_etd4)
}

print(summary(glmm_resp_etd4), corr = FALSE)
```

`Cue_Ali` was removed. 

```{r resp exteded5 model}
file_resp_etd5 <- file.path(folder_lmm, "Resp_lmm_etd5.RData")

# fit the etd5 model
if (!file.exists(file_resp_etd5)) {
  glmm_resp_etd5 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (Cue_C + Sam_C + # Con_C + Ali_C + 
         Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + Cue_Ali + 
         Cue_Con_Sam +  # Cue_Con_Ali +Con_Ali_Sam + Cue_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd5, file = glmm_resp_etd5)
} else {
  load(file_resp_etd5)
}

print(summary(glmm_resp_etd5), corr = FALSE)
```

`Intercept` was removed.

```{r resp exteded6 model}
file_resp_etd6 <- file.path(folder_lmm, "Resp_lmm_etd6.RData")

# fit the etd6 model
if (!file.exists(file_resp_etd6)) {
  glmm_resp_etd6 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (0 + Cue_C + Sam_C + # Con_C + Ali_C + 
         Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + Cue_Ali + 
         Cue_Con_Sam +  # Cue_Con_Ali +Con_Ali_Sam + Cue_Ali_Sam +
         Cue_Con_Ali_Sam | Participant),
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd6, file = glmm_resp_etd6)
} else {
  load(file_resp_etd6)
}

print(summary(glmm_resp_etd6), corr = FALSE)
```

`Cue_C` was removed.

```{r resp exteded7 model}
file_resp_etd7 <- file.path(folder_lmm, "Resp_lmm_etd7.RData")

# fit the etd7 model
if (!file.exists(file_resp_etd7)) {
  glmm_resp_etd7 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (0 + Sam_C + # Con_C + Ali_C + Cue_C +
         Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + Cue_Ali + 
         Cue_Con_Sam + # Cue_Con_Ali +Con_Ali_Sam + Cue_Ali_Sam +
         Cue_Con_Ali_Sam | Participant), # 
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd7, file = glmm_resp_etd7)
} else {
  load(file_resp_etd7)
}

print(summary(glmm_resp_etd7), corr = FALSE)
```

`Cue_Con_Ali_Sam` was removed.

```{r resp exteded8 model}
file_resp_etd8 <- file.path(folder_lmm, "Resp_lmm_etd8.RData")

# fit the etd8 model
if (!file.exists(file_resp_etd8)) {
  glmm_resp_etd8 <- glmer(
    Resp ~ Cue * Congruency * Alignment * SameDifferent + Probability + 
      (0 + Sam_C + # Con_C + Ali_C + Cue_C +
         Cue_Sam + Con_Sam + # Con_Ali + Ali_Sam + Cue_Con + Cue_Ali + 
         Cue_Con_Sam # Cue_Con_Ali +Con_Ali_Sam + Cue_Ali_Sam +
         | Participant), # Cue_Con_Ali_Sam 
    family = binomial(link = "probit"),
    data = df_lmm,
    control = glmerControl(optimizer = "optimx", # calc.derivs = FALSE,
                           optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))
  )
  
  # save(glmm_resp_etd8, file = glmm_resp_etd8)
} else {
  load(file_resp_etd8)
}

print(summary(glmm_resp_etd8), corr = FALSE)
```


### The optimal model

```{r comapre etd and rdc}
# compare the extended and reduced model
anova(glmm_resp_etd8, glmm_resp_rdc2, refit = FALSE)
```

According to BIC, the reduced model (`glmm_resp_rdc`) explained the data better than the extended model (`glmm_resp_etd10`) and, therefore, the reduced model is used as the optimal model.

```{r the optimal model}
glmm_resp_opt <- glmm_resp_rdc2

print(summary(glmm_resp_opt), corr = FALSE)
```


## Estimated marginal means
### Estimated marginal means for hit and false alarm
```{r emm}
emm_resp <- emmeans(glmm_resp_rdc2, ~ Cue + Congruency + Alignment + SameDifferent)

emm_resp

```

### Estimated marginal means for d'
```{r}

emm_d <- contrast(emm_resp, method = "pairwise", simple = "SameDifferent")

emm_d[1:8]

```

#### Composite effects
Composite face effects for top and bottom parts:
```{r}

emm_cf <- contrast(emm_resp, interaction = "pairwise", by = "Cue", infer = TRUE)

emm_cf[1:2]

```

#### Facilitation and interference

```{r}
# Sensitivity d'
emm_fi <- contrast(regrid(emm_resp), interaction = "pairwise", by = c("Cue", "Congruency"), infer=TRUE, adjust = "Tukey")

emm_fi[1:4]
```


```{r}
# Accuracy for same trials only
temp <- contrast(regrid(emm_resp), method = "pairwise", simple = "Alignment", infer = TRUE)
temp[1:4]

# summary(temp[1:2], side = ">")
# summary(temp[3:4], side = "<")

```

```{r}
temp[5:8]
```





